### 2010-10-06 docs/tex/intro.tex
Very little intro

%Introductory Chapter
\chapter{Introduction}
Throughout history mankind has learned from nature. Human aspiration to fly in the sky comes from observing birds in the sky. Aspiration leads to inspiration, and airplanes are invented.