### 2010-12-09 docs/tex/acknowledgment.tex
The acknowledgment page

\begin{center}
\begin{large}
\textbf{Acknowledgement}
\end{large}
\end{center}
I really acknowledge..