### 2010-12-09 docs/tex/title.tex

\begin{titlepage}
\begin{center}

\Large Modeling the Evolution of Mimicry\\[1.5cm]
\large Mohiul Islam\\[2.5cm]
\large A Thesis\\ in\\ The Department\\ of\\ Computer Science\\ and Software Engineering\\[2.5cm]
\large Presented in Partial Fulfillment of the Requirements\\ for the Degree of Master of Computer Science at\\ Concordia University\\ Montreal, Quebec, Canada\\[1.5cm]
\large December 2010\\[2cm]

\vfill

\large \copyright Mohiul Islam, 2010
\end{center}
\end{titlepage}