### 2010-12-09 docs/tex/title.tex

\begin{titlepage}
\begin{center}

\Large \textbf{Modeling the Evolution of Mimicry}\\[1.5cm]
\large \textbf{Mohiul Islam}\\[2.5cm]
\large A Thesis\\ in\\ The Department\\ of\\ Computer Science\\ and Software Engineering\\[2.5cm]
\large Presented in Partial Fulfillment of the Requirements\\ for the Degree of Master of Computer Science at\\ Concordia University\\ Montreal, Quebec, Canada\\[1.5cm]
\large \monthname \ \the \year \\[2cm]
\vfill
\large \copyright \ Mohiul Islam, \the \year
\end{center}
\end{titlepage}