### 2010-12-14 docs/tex/acknowledgment.tex
acknowledgment update

\newpage
\phantomsection
\label{acknowledgments}
\begin{center}
\Large \textbf{Acknowledgments}\\[0.5cm]
\end{center}

My first appreciation goes to my supervisor Dr. Peter Grogono. It is not only a privilege to work with him, but also the inspiration and encouragement that I have received from him during the period of this work, made it possible.

My second appreciation is to my parents Mobinul Islam and Latifa Mobin. Their hard work and sacrifice in life made it possible for me to be able to come all this way.