### 2010-12-22 docs/tex/signature.tex
Signature page updates.

%\newpage
\begin{titlepage}
\begin{center}
\Large \textbf{CONCORDIA UNIVERSITY}\\
\Large \textbf{School of Graduate Studies}\\[1cm]
\end{center}
This is to certify that the thesis prepared\\
\makebox[1.6in][l]{By:} Mohiul Islam\\
\makebox[1.6in][l]{Entitled:}  Modeling the Evolution of Mimicry\\
and submitted in partial fulfillment of the requirements for the degree of
\begin{center}
\small \textbf{Master of Computer Science}\\
\end{center}
complies with the regulations of the University and meets the accepted standards with respect to originality and quality.\\[0.4cm]
Signed by the final examining committee:\\[0.5cm]
\makebox[5cm][l]{} \rule{7cm}{0.1mm} Chair\\ 
\makebox[5cm][l]{} Dr. Eusebius J. Doedel\\[0.5cm]
\makebox[5cm][l]{} \rule{7cm}{0.1mm} Examiner\\ 
\makebox[5cm][l]{} Dr. Thiruvengadam Radhakrishnan\\[0.5cm]
\makebox[5cm][l]{} \rule{7cm}{0.1mm} Examiner\\ 
\makebox[5cm][l]{} Dr. Nawwaf Kharma\\[0.5cm]
\makebox[5cm][l]{} \rule{7cm}{0.1mm} Supervisor\\ 
\makebox[5cm][l]{} Dr. Peter Grogono\\[0.5cm]
\makebox[5cm][l]{Approved by} \rule{10cm}{0.1mm}\\ 
\makebox[5cm][l]{} Chair of Department or Graduate Program Director\\[0.5cm]
\makebox[5cm][l]{} \rule{10cm}{0.1mm}\\ 
\makebox[5cm][l]{} Dr. Robin A. L. Drew, Dean \\
\makebox[5cm][l]{} Faculty of Engineering and Computer Science\\[0.5cm]
\makebox[5cm][l]{Date} \rule{7cm}{0.1mm}

\end{titlepage}