### 2010-12-26 docs/slides/application.tex
New Theme added.

\section{Application}

\frame
{
	\frametitle{Application}
}