### 2010-12-26 docs/slides/conclusion.tex

\section{Conclusion}

\frame
{
	\frametitle{Conclusion}
}