### 2010-12-26 docs/slides/review.tex

\section{Review}

\frame
{
	\frametitle{Review}
}