### 2010-12-26 docs/slides/mimicry.tex
New Theme added.

\section{Mimicry}

\frame
{
	\frametitle{Mimicry}
}