### 2010-12-26 docs/slides/result.tex

\section{Results}

\frame
{
	\frametitle{Results}
}