### 2010-12-27 docs/slides/mimicry.tex
Result added

\section{The Inspiration}

\frame
{
	\frametitle{Mimicry}
}

\subsection{Batesian Mimicry}
\subsection{Mullerian Mimicry}
\subsection{Evolutionary Dynamics}