### 2010-12-27 docs/slides/review.tex
Result added

\section{Review}

\frame
{
	\frametitle{Review}
}

\subsection{Evolutionary Computation}
\subsection{Artificial Life}