### 2010-12-29 docs/slides/conclusion.tex
Analysis and application updated.

\section{Conclusion}

\frame
{
	\frametitle{Conclusion}

	\begin{itemize}
		\item 
	\end{itemize}
}