### 2010-12-29 docs/slides/model.tex
Some slides added.

\section{The Model}

\frame
{
	\frametitle{The Model}
}

\subsection{Past Work}

\frame
{
	\frametitle{Past Work}
}

\subsection{FormAL}

\frame
{
	\frametitle{FormAL Framework}
}

\subsection{Prey}

\frame
{
	\frametitle{Prey}
}

\subsection{Predator}

\frame
{
	\frametitle{Predator}
}