### 2010-12-29 docs/slides/mimicry.tex
Some slides added.

\section{The Inspiration}

\frame
{
	\frametitle{Mimicry}
}

\subsection{Batesian Mimicry}

\frame
{
	\frametitle{Batesian Mimicry}
}

\subsection{Mullerian Mimicry}

\frame
{
	\frametitle{Mullerian Mimicry}
}

\subsection{Evolutionary Dynamics}

\frame
{
	\frametitle{Evolutionary Dynamics}
}

\frame
{
	\frametitle{Mimicry Rings}
}