### 2011-01-02 docs/slides/conclusion.tex
Batesian and Mullerian mimicry slides.

\section{Conclusion}

\frame
{
	\begin{center}
		\LARGE Conclusion
	\end{center}
}

\frame
{
	\frametitle{Conclusion}

	\begin{itemize}
		\item 
	\end{itemize}
}