### 2011-01-10 docs/slides/review.tex
Updates to the review slides.

\section{Review}

\frame
{
	\begin{center}
		\LARGE A Review
	\end{center}
}

\frame
{
	\frametitle{Review}

	\begin{itemize}
		\item Evolutionary Computation
			\begin{itemize}
				\item Genetic Algorithms
				\item Evolutionary Strategies
				\item Evolutionary Programming
				\item Genetic Programming
			\end{itemize}
		\item Artificial Life
			\begin{itemize}
				\item Complex Adaptive System
				\item Echo
			\end{itemize}
	\end{itemize}
}

\subsection{Evolutionary Computation}

\frame
{
	\frametitle{Evolutionary Computation}
	\framesubtitle{History}

%	\begin{itemize}
%		\item Turing (1948): ``Intelligent Machinery", expression ``genetical or evolutionary search"
%		\item Bremermann (1962): Solve optimization problems. Minimize a real-valued function.
%		\item Rechenberg (1964): Evolutionary strategies. Aerodynamic wing design.
%		\item L. Fogel, Owens and Walsh (1965): Merged evolutionary computation and computational intelligence.
%		\item Holland (1975): Genetic Algorithms, Schema theorem.
%		\item Koza (1992): Genetic Programming. Automated method for creating computer program.
%	\end{itemize}

	\begin{itemize}
		\item Turing (1948):
			\begin{itemize}
				\item ``Intelligent Machinery"
				\item expression ``genetical or evolutionary search"
			\end{itemize}
		\item Bremermann (1962):
			\begin{itemize}
				\item Solve optimization problems.
				\item Minimize a real-valued function.
			\end{itemize}
		\item Rechenberg (1964):
			\begin{itemize}
				\item Evolutionary strategies.
				\item Aerodynamic wing design.
			\end{itemize}
		\item L. Fogel, Owens and Walsh (1965):
			\begin{itemize}
				\item Merged evolutionary computation and computational intelligence.
			\end{itemize}
	\end{itemize}	
}

\frame
{
	\frametitle{Evolutionary Computation}
	\framesubtitle{History}

	\begin{itemize}
		\item Holland (1975):
			\begin{itemize}
				\item Genetic Algorithms
				\item Schema theorem.
			\end{itemize}
		\item Koza (1992):
			\begin{itemize}
				\item Genetic Programming.
				\item Automated method for creating computer program.
			\end{itemize}
	\end{itemize}
}

\frame
{
	\frametitle{Evolutionary Algorithm}

	\begin{itemize}
		\item Strategy: survival of the fittest.
		\item Fitness function
		\item Variation operators:
			\begin{itemize}
				\item Mutation
				\item Recombination
			\end{itemize}
		\item Fundamental forces:
			\begin{itemize}
					\item ``Variation operators create diversity; facilitates novelty."
					\item ``Selection acts as a force pushing quality."
			\end{itemize}
	\end{itemize}
}

\frame
{
	\frametitle{Genetic Algorithm}

	\begin{itemize}
		\item Wide set of applications.
		\item Holland's ``Adaptation in natural and artificial systems"
		\item Effective to solve optimization problems.
		\item Ineffective: solution set does not have gnomic representation.
	\end{itemize}
}

\frame
{
	\frametitle{Evolutionary Strategies}

	\begin{itemize}
		\item Self adaptation of strategy parameters.
		\item Real-valued vectors
		\item Recombination: discrete or intermediary.
		\item Mutation: Gaussian perturbation.
		\item Fast and good optimizer for real-valued optimization.
	\end{itemize}
}

\frame
{
	\frametitle{Evolutionary Programming}

	\begin{itemize}
		\item Generate Artificial Intelligence by simulating evolution as a learning process.
		\item Intelligence: System should adapt its behavior to meet specified goal in the environment.
		\item Representation: real-valued vectors.
		\item Parent selection: Deterministic.
		\item Off springs are created via mutation.
		\item No recombination operator.
		\item Mutation: Gaussian perturbation.
		\item Survivor selection: probabilistic.
	\end{itemize}
}

\frame
{
	\frametitle{Genetic Programming}

	\begin{itemize}
		\item Applied to machine learning: prediction and classification.
		\item Behavior: similar to Neural Network.
		\item Very slow process.
		\item Non-linear chromosomes such as trees and graphs.
		\item Mutation is mostly avoided.
		\item Recombination operator work by exchanging trees.
		\item Parent selection: Fitness proportional.
		\item Survivor selection: Replacement of generations.
		\item Difference from genetic algorithm: the chromosomes are trees.
	\end{itemize}
}

\subsection{Artificial Life}

\frame
{
	\frametitle{Artificial Life}
	\framesubtitle{}

	\begin{itemize}
		\item Introduced by Christopher G. Langton.
		\item At `International Conference on the Synthesis and Simulation of Living Systems', Los Alamos National laboratory 1987.
		\item Described by Taylor as a tool for biological inquiry.
	\end{itemize}
}

\frame
{
	\frametitle{Artificial Life}
	\framesubtitle{}

	\begin{itemize}
		\item Wetware system (Molecular level):
			\begin{itemize}
				\item Natural life or derived from it.
				\item Artificial evolutionary process toward the production of RNA molecules.
			\end{itemize}
		\item Software system (Cellular level):
			\begin{itemize}
				\item Initial contribution: John von Neumann.
				\item Self reproducing, computation universal Cellular Automata.
			\end{itemize}
		\item Hardware system (Organism level):
			\begin{itemize}
				\item Organism's sensory and nervous system: body and environment.
				\item Geometric, mechanical, dynamical, thermal properties.
			\end{itemize}
	\end{itemize}
}

\frame
{
	\frametitle{Complex Adaptive System}
	\framesubtitle{}

	\begin{itemize}
		\item Special cases of complex systems.
		\item Diverse and made up of multiple interconnected elements.
		\item Adaptive: capacity to change and learn from experience.
		\item Examples:
		\begin{itemize}
			\item Stock market.
			\item Social insect and ant colonies.
			\item Biosphere and the ecosystem.
			\item Brain and the immune system.
			\item Cell and the developing embryo.
			\item Manufacturing businesses.
			\item Human social group-based endeavor in a cultural and social system.
			\item Large-scale online systems, collaborative tagging or social bookmarking systems.
		\end{itemize}
	\end{itemize}
}

\frame
{
	\frametitle{Complex Adaptive Systems}
	\framesubtitle{Criteria}

	\begin{itemize}
		\item ``Large number of parts undergoing a kaleidoscopic array of simultaneous non linear interactions."
		\item ``Impact in human affairs: aggregate behavior; behavior of the whole."
		\item ``Interactions evolve over time: parts adapt to survive in the environment"
		\item ``Complex adaptive systems anticipate."
	\end{itemize}
}

\frame
{
	\frametitle{Echo}
	\framesubtitle{Criteria}

	\begin{itemize}
		\item Simplicity, thought experiment, primitive internal model.
		\item Fitness: an evolving criteria.
		\item ``The primitive mechanisms in Echo should have ready counter parts in all CAS".
		\item Flexible architecture: incorporate well studied mathematical models.
		\begin{itemize}
			\item Dawkins' biological arms race.
			\item Brower's survival of mimics.
			\item Wicksell's Triangle and overlapping generation models in economics.
			\item Prisoner's dilemma game in political science.
			\item Holland's two-armed bandits in operational research.
			\item Perelson's antigen-body matching in immunology.
		\end{itemize}
		\item Amenable to mathematical analysis.
	\end{itemize}
}