### 2011-01-10 docs/slides/conclusion.tex
Slide updates before submission.

\section{Conclusion}

\frame
{
	\begin{center}
		\LARGE Conclusion
	\end{center}
}

\frame
{
	\frametitle{Conclusion}

	Artificial Life:
	\begin{itemize}
		\item Tool for Biological inquiry
			\begin{itemize}
				\item Success: Proof of Turner's punctuated equilibrium
				\item Proof of Franks and Noble's converge of one large ring.
			\end{itemize}
		\item Nature inspired computer science
			\begin{itemize}
				\item Success: Appropriate emulation of Batesian and Mullerian mimicry.
			\end{itemize}
		\item Solve problem in Computer Science
			\begin{itemize}
				\item Not successful: unable to find the appropriate problem solving scenario.
			\end{itemize}
	\end{itemize}
}

\frame
{
	\frametitle{Conclusion}
	\framesubtitle{Objective of Research}

	\begin{itemize}
		\item Objective of any research:
			\begin{itemize}
				\item Unexplored Territory: find new knowledge, understanding, explanation.
			\end{itemize}
		\item This research:
			\begin{itemize}
				\item New Understanding: Can emulate natures behavior.
				\item Explore a new model.
				\item Better explanation and understanding of nature.
			\end{itemize}
	\end{itemize}
}