### 2011-01-14 docs/slides_shortened/introduction.tex

\section{Introduction}

\frame
{
	\begin{center}
		\LARGE Introduction
	\end{center}
}

\frame
{
	\frametitle{Introduction}

	\begin{itemize}
		\item Mimicry:
			\begin{itemize}
				\item Process of deception.
				\item Evolutionary process.
				\item Organisms survive by deceiving its predators.
				\item Occurs in existence of similar appearing noxious organisms.
				\item Palatable mimics unpalatable for \textit{survival of species}.
			\end{itemize}
		\item Artificial Life:
			\begin{itemize}
				\item Interdisciplinary study of life and \textit{life-like process} using \textit{synthetic method}.
				\item Life-like process: Evolution of mimicry.
				\item Synthetic tools:
					\begin{enumerate}
						\item 3D Visual Environment.
						\item Cellular Automata.
						\item Neural Network.
					\end{enumerate}
			\end{itemize}
	\end{itemize}
}