### 2011-01-16 docs/tex/acknowledgment.tex
Some changes in Acknowledgements.

\newpage
\phantomsection
\label{acknowledgments}
\begin{center}
\Large \textbf{Acknowledgments}\\[0.5cm]
\end{center}

My first appreciation is to my supervisor Dr. Peter Grogono. It is not only a privilege to work with him, but also the inspiration and encouragement that I have received from him during the period of this work, made it possible.

My second appreciation is to my parents Mobinul Islam and Latifa Mobin. Their hard work and sacrifice in life made it possible for me to be able to come all this way.

My third appreciation is to my sister Shagufta Islam, who also made it easier for me to be able to give all the time that I required to complete my thesis.