### 2016-06-28 docs/slides_conference/conclusion.tex

\section{Conclusion}

\frame
{
	\begin{center}
		\LARGE Conclusion
	\end{center}
}

\frame
{
	\frametitle{Conclusion}

	Artificial Life:
	\begin{itemize}
		\item Tool for biological inquiry
			\begin{itemize}
				\item Success: Proof of Turner's punctuated equilibrium
				\item Proof of Franks and Noble's converge of one large ring.
			\end{itemize}
		\item Nature inspired computer science
			\begin{itemize}
				\item Success: Appropriate emulation of Batesian and Mullerian mimicry.
			\end{itemize}
		\item Solve problem in computer science
			\begin{itemize}
				\item Not successful: unable to find the appropriate problem solving scenario.
			\end{itemize}
	\end{itemize}
}